---
title: "615 midterm"
author: "Fangjing Fu"
date: "2016/10/22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(pipeR)
```

# International Food Security

http://www.ers.usda.gov/data-products/international-food-security/

This dataset is the basis for the International Food Security Assessment, 2016-26 released in June 2016.  This annual ERS report projects food availability and access for 76 low- and middle-income countries over a 10-year period.  The dataset includes annual country-level data on area, yield, production, nonfood use, trade, and consumption for grains and root and tuber crops (combined as R&T in the documentation tables), food aid, total value of imports and exports, gross domestic product, and population compiled from a variety of sources. 

* Data File: gfa27.csv
* Variables: 
  \begin{itemize}
  \item  Country: 76 low- and middle-income countries.
  \item  Commodity: Economic Data, Food Aid, Population, Root Crops (R&T), Total Grains/Cereals, Total Grains/Cereals and Root Crops (R&T), Other.
  \item  Item: 24 levels including Area Harvested, Cereal Food Aid, Export Quantity, etc.
  \item  Unit: unit of the data
  \item  Year: 1990-2014
  \item  Amount: data
\end{itemize}


```{r}
#read data
IFS_raw <- read.csv("gfa27.csv")
```

# Tidy data

We selected the variables we're interested in. 

* Variables: 
  \begin{itemize}
  \item  Country: 76 low- and middle-income countries.
  \item  Year: 1990-2014
  \item  cereals_production: Total Grains/Cereals Production Quantity (1000 MT)
  \item  cereals_ava: Total Grains/Cereals Food availability (1000 MT)
  \item  cereals_import: Total Grains/Cereals Import Quantity (1000 MT)
  \item  cereals_export: Total Grains/Cereals Export Quantity (1000 MT)
  \item  cereals_harvested: Total Grains/Cereals Area Harvested (1000 Ha)
  \item  cereals_aid: Cereal Food Aid (Grain Equivalent, 1000 MT)
  \item  total_aid: Total Food Aid (Grain Equivalent, 1000 MT)
  \item  GDP: GDP (constant 2000 US$, million)
  \item  total_export: Exports of goods and services (constant 2000 US$, million)
  \item  total_import: Imports of goods and services (constant 2000 US$, million)
  \item  population: Total Population - Both Sexes (million)
\end{itemize}

```{r}
IFS_tidy <- IFS_raw[,c("Country","Year")] %>% unique()

IFS_tidy <- IFS_raw %>% subset(Commodity=="Population",select=c("Country","Year","Amount")) %>>% full_join(IFS_tidy,., by = c("Country","Year")) 
colnames(IFS_tidy)[3] <- "Population"
```

```{r}
#
tidyIFS <- function(x,y){
  IFS_tidy <- IFS_raw %>% subset(Commodity==x & Item == y,select=c("Country","Year","Amount")) %>>% full_join(IFS_tidy,., by = c("Country","Year"))
  colnames(IFS_tidy)[3] <- y
  return(IFS_tidy)
}
```

```{r}
IFS_tidy <- tidyIFS("Economic Data","Imports of goods and services (constant 2005 US$)")
IFS_tidy <- tidyIFS("Economic Data","Exports of goods and services (constant 2005 US$)")
IFS_tidy <- tidyIFS("Economic Data","GDP (constant 2005 US$)") 
IFS_tidy <- tidyIFS("Food Aid","Total")
IFS_tidy <- tidyIFS("Food Aid","Cereal Food Aid") 
IFS_tidy <- tidyIFS("Total Grains/Cereals","Area Harvested")
IFS_tidy <- tidyIFS("Total Grains/Cereals","Export Quantity") 
IFS_tidy <- tidyIFS("Total Grains/Cereals","Import Quantity")
IFS_tidy <- tidyIFS("Total Grains/Cereals","Food Availability") 
IFS_tidy <- tidyIFS("Total Grains/Cereals","Production Quantity")

names(IFS_tidy) <- c("Country","Year","cereals_production","cereals_ava","cereals_import","cereals_export","cereals_harvested","cereals_aid","total_aid","GDP","total_export","total_import","population")
```

# EDA

##cereals productivity 

###cereals production/cereals availability

```{r}
produc_1 <- IFS_tidy %>% mutate(produc_ava_ratio=cereals_production/cereals_ava) %>% na.omit()

library(ggplot2)

ggplot(produc_1) + aes(x=Country,y=produc_ava_ratio, col=Year) + geom_point() + geom_abline(intercept = mean(produc_1$produc_ava_ratio), slope = 0) + ggtitle("Proportion of Production in Food Availability") + ylab("Ratio") + theme(axis.text.x=element_text(angle=90, hjust=1))

```

###cereals food aid/cereals availability

```{r}
produc_2 <- IFS_tidy %>% mutate(aid_ava_ratio=cereals_aid/cereals_ava) %>% na.omit()

library(ggplot2)

ggplot(produc_2) + aes(x=Country,y=aid_ava_ratio, col=Year) + geom_point() + geom_abline(intercept = mean(produc_2$aid_ava_ratio), slope = 0) + ggtitle("Proportion of Food Aid in Food Availability") + ylab("Ratio") + theme(axis.text.x=element_text(angle=90, hjust=1))

```

###cereals production/cereals area harvested










## Trend of cereals production in 25-year period

```{r}
library(ggvis)

IFS_tidy %>%
  ggvis(~Year,~cereals_production) %>%
  filter(Country %in% eval(input_select(levels(IFS_tidy$Country),label='Country'))) %>% layer_lines()
```

## Cereals Harvested 

```{r}
library(rworldmap)

#Cereals Harvested

cerharv_fun <- function(x){
  cerharv_map <- IFS_tidy %>% subset(Year==x,select=c("Country","cereals_harvested"))
  cerharv_map <- cerharv_map %>% mutate(country_iso3 = unlist(lapply(cerharv_map$Country, rwmGetISO3))) 
  cerharv_map <- joinCountryData2Map(cerharv_map, joinCode = "ISO3", nameJoinColumn = "country_iso3")
  return(cerharv_map)
}



```

```{r}
#Cereals Harvested in Africa

```


